Notes and Problems from My Research\vspace{-1ex}

### 2019-03-12 Hopf Algebra Notes/Hopf Algebras.tex
Fortnightly Update

Returning to the example from the beginning of class (\ref{ex-22519}), if $V$ acts on $A$,
then 
\begin{itemize}
	\item $g\in \Aut_{alg}(A)$
	\item $x\in\Der_g(A)$
\end{itemize}
In particular, notice that $x(1)=0$ and $x(ab)=x(a)b+g(a)x(b)$.
\end{document}

>>>>

Returning to the example from the beginning of class (\ref{ex-22519}), if $V$ acts on $A$,
then 
\begin{itemize}
	\item $g\in \Aut_{alg}(A)$
	\item $x\in\Der_g(A)$
\end{itemize}
In particular, notice that $x(1)=0$ and $x(ab)=x(a)b+g(a)x(b)$.

\section{March 11, 2019}
We took a week off since James was in Germany for a conference. We're coming back together for this
last week of the quarter.

Here is one result that was brought up at the conference. It was proved by a grad student and might 
have some missing hypotheses:
\begin{thm}
	If $\ch k=0$ (and some other mild hypotheses) then the Etingof-Ostrik conjecture holds for $H$ 
	if and only if it holds for $H^*$. That is,
	\[\operatorname{kdim}\oplus \Ext_H^i(k,k)=\operatorname{kdim}\oplus \Ext^i_{H^*}(k,k)\]
	where $\operatorname{kdim}$ means Krull dimension.
\end{thm}

\subsection{Example of the Day}
\begin{ex}[Small Quantum Group of $\sl_2$]
	Let $q\in k$ be a primitive $l^{th}$ root of unity ($l\ge 2$). Then
	\[u_q(\sl_2)=k\langle K,K^{-1},E,F\rangle/R\]
	where $R$ is the ideal for the relations
	\begin{align*}
		KK^{-1}=K^{-1}K=1\\
		KEK^{-1}=q^2E\\
		KFK^{-1}=q^{-2}F\\
		EF-FE=\frac{K-K^{-1}}{q-q^{-1}}\\
		E^l=F^l=K^0\\
		K^l=1
	\end{align*}
	Now $\dim_k u_q(\sl2)=l^3$ and this is (symmetric) Frobenius and noncommutative and noncocommutative.
	As a coalgebra, $K$ is grouplike and $E,F$ are almost primitive:
	\[\Delta E=E\otimes 1+K\otimes E,\quad \Delta F = F\times K^{-1}+1\otimes F\]
	and $S(E)=-K^{-1}E$ and $S(F)=-FK$.

	Then $\sl_2=ke\oplus kf\oplus kh$ and there is a map $e\mapsto E$, $f\mapsto F$ and $K=\exp(h)$. Hmm
\end{ex}

\subsection{Last Lecture}
I wasn't here, so I missed this part. Let $A$ be a Noetherian connected graded AS-regular algebra and let
$V$ be a Hopf algebra action on $A$ such that each $A_i$ is a left $V$-module.
\begin{enumerate}
	\item $k:=A/A_{\ge 1}$ and $\oplus_0^\infty \Ext_A^i(k,k)$ is an algebra.
	\item (S. Paul Smith's Result) $\oplus \Ext_A^i(k,k)$ is Frobenius and $\Ext_A^d(k,k)$ is one
	dimensional over $k$ when $d=\gldim A$.
	\item  Each $\Ext^i_A(k,k)$ is a left $V$ module.
	\item $(\oplus \Ext_A^i(k,k))^{op}$ is a left $V$ algebra.
\end{enumerate}

Now pick any nonzero element $e\in \Ext_A^d(k,k)$ where $d$ is as above. For all $v\in V$, 
$v\cdot e\in \Ext_A^d(k,k)$ and we write $v\cdot e=\operatorname{hdet}(v)v$ for $\operatorname{hdet}(v)\in k$.
\begin{defn}
	A map $\operatorname{hdet}:V\to k$ that fits in as above is called the \textbf{homological determinant}
	of the $V$-action on $A$.
\end{defn}
\begin{lem}
	$\operatorname{hdet}$ is an algebra map.
\end{lem}
\begin{lem}
	If $A=k[x_1,\dots,x_n]$, $V=kG$, $G\subseteq GL(n)$ then
	\[\operatorname{hdet} g = \det g\]
	for all $g\in G$.
\end{lem}
\begin{defn}
	We say that $\operatorname{hdet}$ is trivial if $\operatorname{hdet}(v)=\varepsilon(v)$ for all $v\in V$.
\end{defn}

Recall the theorem
\begin{thm}[Watanabe]
	Suppose $G$ is a finite subgroup in $GL_n$. If $\deg g=1$ for all $g\in G$, then $k[x_1,\dots,x_n]^G$ is Gorenstein.
\end{thm}
\begin{thm}[Noncommutative Watanabe]
	Let $A$ be a Noetherian connected graded AS-regular algebra. Let $V$ be a semisimple Hopf algebra acting on 
	$A$ such that each $A)i$ is a left $V$-module. If $\operatorname{hdet}$ is trivial, then $A^V$ is AS Gorenstein.
\end{thm}
\begin{rmk}
	This last result is important because it distinguishes between the commutative and noncommutative cases: in the commutative
	case the only AS-regular ring is the polynomial ring. This shows us that there is in fact a vast class of them 
	in the noncommutative case. In fact it's even bigger!
\end{rmk}
Recall that $g\in GL_n$ is a reflection if $g$ is similar to $\operatorname{diag}(1,\dots,1,\sqrt{-1})$. Let $R$ be the subgroup of $G$ generated
by the reflections in $G$. 
\begin{thm}[Watanabe II]
	Suppose $G$ is a finite subgroup of $GL_n$. Then $k[x_1,\dots,x_n]^G$ is Gorenstein if and only if
	$\det \bar g=1$ for all $g\in G/R$.
\end{thm}

\subsection{Noncommutative Reflections}
\begin{defn}
	Let $g\in\Aut_{gralg}(A)$. Then
	\[\operatorname{Tr}(g)=\sum_0^\infty \operatorname{tr}(g|_{A_i})t^i\in k[[t]].\]
\end{defn}
Then a result (which?) gets us that $\operatorname{Tr}(g)=\frac{1}{p(t)}$ for a polynomial $p$.
Consider $\tilde A=k_{p_{ij}}[x_1,\dots,x_n]$ where $x_jx_i=p_{ij}x_ix_j$ for all $i<j$. Then 
\begin{defn}
	$g\in\Aut_{gralg}(\tilde A)$ is a reflection if $\operatorname{Tr}(g)=\frac{1}{(1-t)^{n-1}(1-\lambda t)}$.
\end{defn}
\end{document}

### 2019-03-12 Lie Algebra Notes/Lie Algebras.tex
Fortnightly Update

%characteristic
\DeclareMathOperator{\ch}{char}

@@

\section{February 28, 2019}

\end{document}

>>>>

%characteristic
%\DeclareMathOperator{\char}{char}
\DeclareMathOperator{\ch}{ch}

@@

\section{February 28, 2019}
Had an appointment. :(

\section{March 4, 2019}
Notice that we should read sections 21 and 24 of the text, but not 22 or 23. There are notoriously hard sections and it will end 
up that we don't actually need 23, using the appendix to section 24.

Last itme we identitified the highest weight of the definitng representation $V=k^n$ of any classical Lie algebra
$L\subseteq \gl_n(k)$. The highest weight is $e_1$ in essentially all cases of type $A_{n-1}$ -- modify to
($\frac{n-1}{n},\frac{-1}{n},\dots, \frac{-1}{n}$).

\subsection{Creating lager representations}
We have see the tensor product $\otimes V_i$ of representations $V_i$ over a Lie algebra $L$ that has a Lie sructure defined by 
\[x(v_1\otimes v_n)=\sum_{i=1}^nv_1\otimes\cdots\otimes xv_i\otimes\cdots\otimes v_n.\]

Certain quotients of $T^nV$, the $n$-fold tensor product, also inherit an $L$ module structure. Specifically, $S^nV=T^nV$ modulo
the ideal generated by $v\otimes w-w\otimes v$ intersected with $T^nV$. These are the homogeneous polynomials of degree $n$ on a basis of $V$
as an $L$-module. Similarly one can corrects the exterior power $\Lambda^nV$.

Recall that $S^nV$ grows in dimension while the exterior power grows like binomial coefficients.

In particular, any classical Lie algebra $L\subseteq\gl_n(k)$ acts on $S^mk^n$ and $\Lambda^mk^n$. Luckily
there is one phrase that sums up how $L$ acts:

\textit{All summetric and exterior powers of $k^n$ in this situation are irreducible, modulo the obvious relations.}

\begin{itemize}
	\item For $A_{n-1}$ ($\sl_n(k)$), all powers $S^mk^n$ and $\Lambda^mk^n$ are irreducible.
	\item For $B_n$ and $D_n$, recall that the corresponding group $\so_n(k)$preserves a symmetric bilinear nondegenerate
	form, which may be thought of as an element of $S^2k^n$. Thus $S^mk^n$ contains a copy of $S^{m-2}k^n$ and then
	$S^mk^n/S^{m-2}k^n$ is irreducible under the $L$-action, identifying $S^{m-2}k^n$ with its copy in $S^mk^n$. the highest weight is $me.$
	\item For type $C$, the exterior power $\Lambda^mk^n$, as long as $2\le m\le\frac{n}{2}$. Then the underling group $\sp_n(k)$ preserves
	a skew-symmetric nondegenerate bilinear form, whence $\Lambda^mk^n$ contains acopy of $\Lambda^{m-2}k^n$ and the corresponding 
	quotient is irreducible. The highest weight is $m$ 1's followed by $n-m$ zeros.
\end{itemize}

More generally, there is a purely multilinear algebraic construction starting from $V=k^n$ which produces all irreducible finite dimensional representations in 
types $A$ and $C$. For $B$ and $D$, we need an additional ingredient of the half-splin representations seen in hw.

\subsection{A Road Map}
We now head towards a road map of all finite dimensional irreducible represetnations of a semisimple Lie algebra $L$. That is, recall 
that every such representation $V$ is the sum $\oplus V_\mu$ of its weight spaces, each occuring with some multiplicity $m_\mu=\dim V_\mu$.

We want an implicite formulat for all the multiplicities. We introduce some formalisms:
\begin{defn}
	$\ch V$ is the formal character of $V$ and is defined to be
	\[\ch V=\oplus_\mu m_\mu e^\mu\]
	where the sum is over all weights of $V$ (the weight lattice).
\end{defn}

We add any two finite formal sums in the obvious way and multiplication is also obvious in that $e^\mu s^\sigma=e^{\mu+\sigma}$.

Then we can actually take the formal characters $\ch M_\lambda$ of \textbf{Verma modules} $M_\lambda :=\mathcal U/(L_\alpha,h-\lambda(h),\alpha\in\Phi^+)$
where $\ch M_\lambda=\sum a_\mu e^\mu$ where (as long as $\lambda$ is a weight) all $\mu$ occuring will lie in the same lattice.

We will work out an expression for $\ch L_\lambda$, the formal character of the irreducible highest weight module of heighest weight $\lambda$, a dominant integral weight $\lambda$
in terms of various $M_\mu$. We will see that $\ch L_\lambda$ will be a quotient of two simple alternating sums of ${e^\mu}'s$. To do this, we will have to dive into the structure of $M_\lambda$
for a general $\lambda$.

Notice that any subquotient (quotient of a submodule) of $M_\lambda$ has a weight that is maximal in the partial order, so also a highest weight vector of that highest weight. 
Then that vector generates a highest weight module with a unique irreducible quotient $L_\mu$ for some $\mu$ so every irreducible subquotient 
of $M_\lambda$ takes the form $L_\mu$ for some $\mu$.

Now only finitely many $L_\mu$ can occur as subquotients of $M_\lambda$ for a fixed $\lambda$. 
To prove theis, we need to revisit and generalize the idea of a \textbf{Kasimir element.} Recall the Killing form,
which is nondegenerate on $L$. Set up dual basis of this form so that $\kappa(x_\alpha,z_\alpha)=1$ for $x_\alpha\in L_\alpha$ and $z\in L_{-\alpha}$.

Set
\[c=\sum_{\alpha\in\Phi^+}(x_\alpha z_\alpha+z_\alpha x_\alpha)+\sum_{i=1}^nh_i^2\]
and notice that $c$ lies in the center of $\mathcal U$ -- it commutes with $L$. Then consider 
(for $\lambda$ a highest weight vector)
\[c v_\lambda=\sum_1^r\lambda(h_i)^2+\sum_{\alpha\in\Phi^+}(z_\alpha x_\alpha+[x_\alpha,z_\alpha]+z_\alpha x_\alpha)v_\lambda=(\lambda\cdot \lambda+\lambda(2\rho))v_\lambda\]
where $2\rho$ is the sum of the positive roots (called the ``rho shift'')

\section{March 6, 2019}
Last time we stated that in the classical case, if $V=k^n$ is the defining (or natural) representation of
$L\in\gl_n(k)$, then the symmetric and exterior owers are irreducible under $L$, except for the ``obvious relations.''

The obvious relations were that, if $L$ is orthogonal, $S^2V$ has an $L$-invariant vector (sent to zero)
so $S^mV$ is not irreducible for $m\ge 2$, but $\Lambda^mV$ is irreducible of highest weight $(1,\cdots,1,0\cdots,0)$  (with $m$ ones)
for $m\le\lfloor\frac{n}{2}\rfloor$ where $n=\dim V$.

Likewise if $L$ is symplectic, the $S^mV$ are all irreducible with highest weight $(m,0,\dots,0)$.

\subsection{Return to (Formal) Characters}
Recall that last time we studyies the Verma modules $M_\lambda=\mathcal U/(L_\alpha, h-\lambda(h): h\in H, \alpha>0)$.
These all have irreducible quotients $L_\lambda$. SO we introduced the formal character $\ch M_\lambda$ of $M_\lambda$
to be $\ch M_\lambda=\sum n_\mu e^\mu$ where $|mu$ runds over the $H$ weights of $M_\lambda$ 
and $n_\mu=(M_\lambda)_\mu$, the dimension of the corresponding weight space.

\begin{rmk}
	We'll do a quick aside to explain the connection between the character of a group representation vs 
	one of a Lie algebra. Recall that in the group case, we are just taking the determinant of a matrix 
	(the image of $g$ under the representation).

	In general, there is an exponential map from any Lie algebra to a Lie group $G$ with Lie algebra $L\subseteq \gl_n(k)$.
	Here $G=GL(n,k)$ and $\exp \mu=\sum_{i=0}^\infty \frac{\mu^i}{i!}$ (we can work here over $\C$ so the series 
	always converges.)

	Then if $h\in H$, the Cartan subalgebra and if $N$ is a finidte simensional subquotient, then
	\[\tr e^h)=\sum n_\mu e^{\mu(h)}.\]
	we are extending to the case of infinite dimensional $M_\lambda$ without worrying about convergence.

	We saw last time that any $M_\lambda$ admits a finite composition series $0=M_0\subseteq\cdots\subseteq M_m=M_\lambda$
	where successive quotients are irreducible and each is isomorphic to some $L_\mu$. Then we are saying
	\[\ch M_\lambda=\sum_{1}^m\ch(M_i/M_{i-1}).\]
\end{rmk}

Let's specialize this discussion to the case when $\lambda$ is \textbf{dominant integral}, which is the case of 
greatest interest, so that $L_\lambda$ is finite-dimensional. Now
\[\ch L_\lambda=\sum_\mu m_\mu\ch M_\mu\]
and we will work out exactly which $\ch M_\mu$ occur and the $m_\mu$ in each case. More precisely,
\[\ch M_\mu=e^\mu\prod_{\alpha\in\Phi^+}(1+e^{-\alpha}+e^{-2\alpha}+\cdots)=e^\mu\prod_{\alpha\in\Phi^+}\frac{1}{1-e^{-\alpha}}=\frac{e^{\mu+\rho}}{\prod(e^{\alpha/2}-e^{-\alpha/2})}\]

Now if $|lambda$ is dominant integral we know that $\dim(L_\lambda)_\mu=\dim(L_\lambda)_{w\mu}$
for any $w\in W$. So $w(\ch L_\lambda)=\ch L_\lambda$. Now applying any simple reflection $s_\alpha$ to the product above
just premutes the terms except the one including $\alpha$, which is sent to its negative. Thus if $\Delta$ is 
the denominator, $s_\alpha\Delta=-\Delta.$

Hence $w\Delta=(\det w)\Delta$ where $\det w$ is computed with respect to the $W$ action on $H^*$. If 
$\ch L_\lambda=\sum m_\mu \ch M_\mu=(\sum m_\mu e^{\mu+\rho})/\Delta$, it must be the case that 
$w(\sum m_\mu e^{\mu+\rho})=(\det w)\sum m_\mu e^{\mu+\rho}$. THus we get that the coefficent of 
$e^{w\lambda+\rho}$ in the sum must be $\det w$. But no other coefficients can appear -- $m_\mu=0$
if $\mu\ne w(\lambda+\rho)-\rho$ for if any $m_\mu\ne -$ then $m_{w\mu}\ne 0$ for some
\textbf{dominant} $w\mu$ lying below $|lambda$< but $w\mu+\rho$ would have the same length as $\lambda+\rho$.

This leads us to
\begin{thm}
	The Weyl Charcter formula:
	\[\ch L_\lambda=\frac{\sum_{w\in W}(\det w)e^{w(\lambda+\rho)}}{\prod_{\alpha\in \Phi^+(e^{\alpha/2}-e^{-\alpha/2}}}.\]
\end{thm}
\begin{rmk}
	If $\lambda=0$ then $L_\lambda$ is the trivial representation having weight zero with multiplicity one and no other weight. So 
	\[\prod(e^{\alpha/2}-e^{-\alpha/2}=\sum_W(\det w)e^{w\rho}\]
	so 
	\[\ch L_\lambda=\frac{\sum_{w\in W}(\det w)e^{w(\lambda+\rho)}}{\sum_{w\in W}(\det w)e^{w\rho}}\]

	There is one small annoyance: that there is no product expression for the nmerator, except for the case when $\lambda=n\rho$.
	Then we see that the weights of $L_n\rho$, with heighest weight $n\rho$ are of the form $n\rho-\sum n_\alpha\alpha$ for positive roots $\alpha$

	The multiplicitiy of any weight is just the number of ways it can be written in this form.
\end{rmk}

\section{March 8, 2019}
We will begin by mopping up some minor technical issues in the Weyl character formula from last time.

We have Kostant's multiplicity formula: in computing $\ch L_\lambda$, we used that it is the sum over $w\in W$
of $(\det w)\ch M_{w(\lambda+\rho)-\rho}$, then $m_\mu$, the multiplicity of $\mu$ in $L_\lambda$ is 
\[m_\mu=\sum_{w\in W}(\det w)P(w(\lambda+\rho)-(\mu+\rho))\]
where $P(\nu)$ is the number of ways to write $\nu$ as nonnegative ineger combination of positive roots.

\subsection{Computing Dimension of $L_\lambda$}
We use the Weyl character formula to compute a finite dimensional representation of highest weight $\lambda$, assumed dominant integral.
Start with $R$, the rings of $\Z$ combinations of $e^\mu$ for $\mu$ in the root lattice. We have a homomorphism $v:R\to \Z$ taking $e^\mu\to 1.$
The basic idea here is to apply this to the top and bottom of the Weyl character formula. We get zero in both
cases. So we use something like L'H\^opital's rule.

More precisely, given any positive root $\alpha$, there is a derivation $\delta_\alpha$ of $R$ sending $e^\mu$ to $(\mu \alpha)e^\mu$ (or something similar).
Let $\delta=\prod_{\alpha>0}\delta_\alpha$ , which is not a derivation. But we can compute $\delta$
on any product, term by term:
\[\Delta(\ch L_\lambda)=\sum_{w\in W}(\det w)e^{w(\lambda+\rho)}\]
where $\Delta$ is the Weyl denominator from before. Apply $v\delta$ to each side.

On the left side, the nly remaining terms are from $v(\delta\Delta)\ch L_\lambda$ since $v(\Delta)=0$.
On the other hand when we apply $\delta$ to a single $(\det w) e^{w(\lambda+\rho)}$, you get the 
same result, independent of $w$ since any $w$ is the product of simple reflections and 
\[\det(ws_\alpha)=\det w\prod_{\beta>0}(w(\lambda+\rho)\cdot \beta)=-\det w\prod_{\beta>0}(ws_\alpha(\lambda+\rho)\cdot \beta)\]

With a little more computation we get 
\begin{thm}
	\textbf{Weyl's Dimension Formula}
	\[\dim L_\lambda =\frac{\prod_{\beta>0}(\lambda+\rho)\cdot\beta)}{\prod_{\beta>0}\rho\cdot \beta}\]
\end{thm}

In general when applying this formula, we usually multiply each term $w(\lambda+\rho)\cdot\beta$ or $\rho\cdot \beta$ by 2, 
divide by $\beta\cdot \beta$ so all terms on top and bottom are integers.

\subsection{What this means}
Recall that for any root system $\Phi$ there is a dual root system $\Phi^v=\{2\alpha/(\alpha\cdot\alpha)|\alpha\in\Phi\}$
which is again a root system with a positive rootsystem going to a positive roots system and a simple subsystem going to a simple subsystem,

Applying this formula to $G_2$, notice that the dual system is isomorphic to $G-2$, but the duality
interchanges long and short roots. The simple roots are $\alpha=(0,1,-1)$ and $\beta=(1,-2,1)$
and the positive roots are $\alpha,\beta,\alpha+\beta,\beta+2\alpha,\beta+3\alpha,2\beta+3\alpha.$

In the dual system, we get simple roots the same thing but with $\alpha,\beta$ swapped for $\beta^v$ and $\alpha^v$, respectively. THen applying the formula 
with $\lambda a^v=1$ and $\lambda\beta^v=0$, we get $\dim L_\lambda=7$. Thus $G_2$ has a seven dimensional representation. This tells us that a Lie algebra $L$ 
of type $G_2$ embeds in $\gl_7$, and in fact in $\so_7$. So $L$ may be realized as a subalgebra of $\so_7$ (which is type $B_3$)
and you can see these computations explicitly in $\S 19$ of the book.

If instead we have $\lambda\cdot a^v=0$ and $\lambda\cdot \beta^v=1$, then $L_\lambda$ is 
14 dimensional, and this is the adjoint representation of $G_2$. The largest exception simple Lie algebra s of type $E_8$
and it is the unique simple Lie algebra $L$ such that $L$ itself is the unique nontrivial $L$-module of smallest dimension (which is 248).

\subsection{Next Week}
We are going to talk about simple Lie algebras over $\R$ -- we will find that it will somehow 
mirror the classification we mentioned for Lie groups (where they were parameterized by subgroups between the weight
and integer lattices). This time we will find that we get two types: complex simple Lie algebras which are just 
considered to be real and real forms of complex simple lLie algebras such that $L\otimes_R\C$ is simple.

\section{March 11, 2019}
Today we are going to say something about semisimple Lie algebras over non algebraically closed fields (in particular $\R$).
In the first week we defined the classical Lie algebras and gave bases for them. We say the commutator of any two basis vectors was an 
integer combination of the the others.

Now let $k$ be a field of characteristic zero (this isn't needed everywhere so we will point out where) and let $\Phi$
be a root system. We have already seen the construction of an abstract semisimple Lie algebra over $k$ with root system $\Phi$
by generators and relations. We didn't need anything about the field except that we needed formulas (power series)
to implement the action of the Weyl group, which involves denominators. In fact, we only get twos and threes.

This doesn't work over $\Z$, but let $L$ be the complex semisimple Lie algebra with root system $\Phi$. Let $\sigma$ be a
Chevalley automorphism of $L$ (recall it acts on a Cartan subalgebra by $-1$ and interchanges the root spaces $L_{\alpha}$ and $L_{-\alpha}$).
Now for all $\alpha\in\Phi^+,$ pick $0\ne x_\alpha\in L_\alpha$. We know that $[x_\alpha,-\sigma(x_\alpha)]$ is a multiple
of $h_\alpha(w(2\alpha)/(\alpha,\alpha))$. If we replace $x_\alpha$ by $cx_\alpha$ for any $c\in\C$,
then $-\sigma(x_\alpha)$ is also replaced by $-c\sigma(x_\alpha)$, so $[x_\alpha,-\sigma(x_\alpha)]$ is replaced by $c^2$
times itself. Thus we can choose $c$ to scale $x_\alpha$ such that htis bracket is exactly $h_\alpha$.

Choose a basis of $L$ as follows: $x_\alpha\in L_\alpha$ for $\alpha>0$ and $-\sigma(x_\alpha)\in L_{-\alpha}$ for $\alpha>0$. 
Then pick $h_i=h_{\alpha_i}$ for all the simple roots $\alpha_i$. THen 
\[[x_\alpha,x_{-\alpha}]=h_\alpha\quad [h_i,x_{\alpha}]=\alpha(\alpha_i)h_\alpha\in\Z h_\alpha\quad [x_\alpha x_\beta]=c_{\alpha\beta}x_{\alpha+\beta}\]
and by applying $\sigma$, we get that $[x_{-\alpha}x_{-\beta}]=c_{-\alpha-\beta}$, so $c_{-\alpha-\beta}=-c_{\alpha\beta}$

Then by thm 25.2(a) in the text, this implies $c_{\alpha\beta}=\pm(r+1)$ where $r$ is the largest integer 
such that $\beta-r\alpha\in\Phi$. The upshot here is that any bracket of basis vectors is an integer combination of basis vectors.

So we see that htere is a Lie algebra (usually semisimple) over any basefield $k$ with root system $\Phi$, called \textbf{split}
(having a root space decomposition and a lage abelian subalgebra consisting of semisimple elements). We now want to look at semisimple Lie algebras 
\textit{not} admitting root space decompositions -- the main example is $\R$.

Let $\Phi$ be a root system, $L$ the corresponding complex semisimple Lie algebra, and $\{h_i,x_\alpha\}$
a Chevalley basis. Look at the $\R$-subalgebra $L'$ of $L$ spanned by $x_\alpha-x_{-\alpha}$ and $i(x_\alpha+x_{-\alpha})$ as well as $ih_1,\dots,ih_l,$ for all $\alpha\in\Phi^+$.
Now $L'$ is closed under the bracket and the Killing form $\kappa$ on $L'$ is negative definite.

But $L'$ doesn't admit a root space decomposition because $L'$ is toral -- it consists \textit{entirely} of semisimple elements:
possible since in fact the semisimple elements are \textbf{not} diagonalizable (in a non-algebraically-closed field semisimple and diagonalizable don't have to coincide)
and in fact, they will all have purely imaginariy eigenvalues.

We call $L'$ \textbf{compact} -- note that it is not compact as a topological space, but its adjoint group $\Int L$, generated 
by $\exp\ad x$, is a compact Lie group. It is, in fact, a closed subgroup of $O(n,\R)$ (a compact group)
and $n=\dim_\R L'=\dim_\C L$, Since it preserves $\kappa$, a negative definite symmetric bilinear form, the \textbf{compact real form of $L$.}

\subsection{Unitary matrices}
An astute reader will notice that we haven't mentioned the Lie groups formed by unitary matrices, for instance (and 
serving as a guide to understand the rest) $\mathfrak{su}(n,\C)$, the set of all $n\times n$ complex matrices that are skew-adjoint
with respect to the Hermetian form $((v_i),(w_i))=\sum_1^n v_i\bar w_i$ and of trace zero.
Notice that although these matrices are complex, the algebra is not stable under complex multiplication, so it is not a \textit{complex} Lie algebra.
The correspoinding group $SU(n,\C)$ consists of all $g\in GL_n(k)$ such that $(gv,gw)=(v,w)$ for all $v,w\in\C^n$.

\subsection{Orthogonal matrices}
$\mathfrak{so}_n(\R)=\mathfrak{o}_n(\R)$, the skew-adjoint matrices with respect to the dot product on $\R^n$ has correspinding group $SO(n,\R)$.
The lie algebra is all the $n\times n$ skey-symmetric matrices in the ordinary sense.

Over $\R$, the group that we would get from our original construction is not $SL(n,\R)$, but something else.

\subsection{Symplectic Case}
Starting with $\sp_{2n}(\C)$, take the intersection with $\mathfrak{su}_{2n}(\C)$ -- this is still a real Lie algebra.
We can think of this as right-linear transformations of $\mathbb{H}^n$ (quaternions) preserving a skew-Hermitian form:
\[((v_1,\dots,v_n)(w_1,\dots,w_n))=\sum_1^nv_ij\bar w_i.\]

Notice that we say ``right-linear'' because $\mathbb{H}^n$ admits (different) left- and right-vector space structures.

Then the group in question is the symmetry group of this form, the Lie algebra consists of matrices skew-adjoint with respect to it.

\subsection{Other Real Forms}
The above concludes our discussion of the \textit{compact} Lie algebra over $\R$, but there are even 
more real forms that are neither compact nor split. There are only finitely many of these, however, and each is partly compact and partly not.
These are of the form $L=K+P$ where $K$ is compact and $[PP]\subseteq K$. All the elements in $K$ will be semisimple
with purely imaginary eigenvalues and $P$ will consist of both semisimples and nilpotents, but the semisimples will have 
real eigenvalues.

One such example is defined by the form 
\[(v_1,\dots,v_n)(w_1,\dots,w_n)=\sum_1^m v_i\bar w_i-\sum_{m+1}^n v_j\bar w_j.\]
This is called the \textbf{symmetric group} $U(n,n-m)$.

%TODO: FIX CHAR/CH
\end{document}

### 2019-03-12 Reading Notes/reading_notes.tex
Fortnightly Update

%schemes
\newcommand{\Ga}{\ensuremath\mathbf{G_a}}
\newcommand{\Gm}{\ensuremath\mathbf{G_m}}
\newcommand*{\GL}[1]{\ensuremath\mathbf{GL_#1}}
\newcommand*{\SL}[1]{\ensuremath\mathbf{SL_#1}}
\newcommand*{\mun}[1]{\ensuremath\mathbf{\bm\upmu_#1}}
\newcommand*{\alp}[1]{\ensuremath\mathbf{\bm\upalpha_#1}}

@@

>>>>

%schemes
\newcommand*{\bbG}{\ensuremath\mathbb{G}}
\newcommand*{\Ga}{\ensuremath\mathbf{G_a}}
\newcommand*{\Gm}{\ensuremath\mathbf{G_m}}
\newcommand*{\GL}[1]{\ensuremath\mathbf{GL_#1}}
\newcommand*{\SL}[1]{\ensuremath\mathbf{SL_#1}}
\newcommand*{\mun}[1]{\ensuremath\mathbf{\bm\upmu_#1}}
\newcommand*{\alp}[1]{\ensuremath\mathbf{\bm\upalpha_#1}}

@@

\newpage
\section{Julia \& Dave Benson's preprint}
As an aside, in response to me saying I needed to practice doing computations with spectral sequences,
Julia recommended I read her current work \cite{julia-dave} with Dave Benson wherein there is a supposedly
straightforward spectral computation in support of their larger work \cite{detecting-nilpotence}.

I have already spent some time reading this paper, but the core objects in question are \textbf{supergroup schemes}
which have coordinate Hopf algebras algebras with a $\bbZ/2$ grading that are furthermore (graded) cocommutative (just 
as group schemes correspond to finite dimensional cocommutative Hopf algebras). The fundamental example is $\bbG_a^-$,
which has as its coordinate ring $k[x]/(x^2)$, or equivalently the exterior algebra on a single element $x$,
where $x$ is taken to be a primitive element.

Then if we let $A$ be the coordinate algebra of $\bbG_a^-\times\bbG_a^-$ and $H$ that of $\bbG_{a(r)}\times(\bbZ/p)^s$ where $\bbG_{a(r)}$ is the
$r^{th}$ Frobenius kernel of $\bbG_a$, we can define the \textbf{smash product} or in this case, since 
$A$ is graded cocommutative, the \textbf{semidirect product} $A\#H=A\rtimes H$ where we define the action
to be (if $\tau:H\otimes A\to A$ is the action of $H$ on $A$)
\[(a\otimes h)\cdot(a'\otimes h')=\sum (-1)^{|h_1||b|}a\tau(h_1\otimes b)\otimes h_2g.\]

So we use that $A\cong \Lambda(k[u,v])$, the exterior algebra on two (primitive) generators $u$ and $v$
and that, if $t_i$ is the generator for the coordinate algebra of the $i^{th}$ copy of $\bbZ/p$, and if 
we let $g_i=t_i-1$, these along with the $s_1,\dots s_r$ generating the Frobenius kernel, generate $H$.
